---
title: "Cyclistic Bike Share"
format: pdf
editor: visual
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(scales)
library(leaflet)
```

## Kaggle Source: 

```{r}
#| echo: false
rm(list = ls())
# Get a list of all files in the data directory
files <- list.files("data", full.names = TRUE)

# Read and combine all files into a single dataframe
bike_data <- files %>%
  map_dfr(read_csv) %>%
  filter(!is.na(start_station_name)) %>% select(-ride_id,-start_station_id,-end_station_id)

```


## Data Wrangling

```{r}
#| echo: false
bike_data <- bike_data %>%
  mutate(date = as.Date(started_at),
         day_of_week = wday(date, label = TRUE),
         month = month(date, label = TRUE),
         hour = hour(started_at))

# Calculate trip duration by subtracting started_at from ended_at
bike_data <- bike_data %>%
  mutate(trip_duration = as.numeric(ended_at - started_at, units = "mins"))  %>%
  mutate(trip_duration = round(trip_duration)) %>% filter(trip_duration >1)
```

## Export clean data

```{r}
# write_csv(bike_data,file="bike_data.csv")

```


```{r}
#| echo: false
bike_data$rideable_type <- as.factor(bike_data$rideable_type)
bike_data$member_casual <- as.factor(bike_data$member_casual)
```
```{r}
bike_data %>% select(rideable_type,member_casual,day_of_week,month) %>% summary()
```

## A histogram of trip durations

```{r}
bike_data %>%
  ggplot(aes(x = log10(trip_duration))) +
  geom_histogram() +
  labs(title = "Histogram of Trip Durations",
       x = "Trip Duration (mins)",
       y = "Count")
```

## A bar chart of customer types

```{r}
bike_data %>%
  ggplot(aes(x = member_casual, fill = member_casual)) +
  geom_bar() +
  labs(title = "Bar Chart of Customer Types",
       x = "Customer Type",
       y = "Count")
```
##  A line chart of bike usage by month

```{r}
bike_data %>%
  group_by(month) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = month, y = count, fill = month)) +
  geom_col() +
  labs(title = "Column Chart of Bike Usage by Month",
       x = NULL,
       y = "Count") +
  theme(axis.text.x = element_blank())
```
## A boxplot of trip durations by customer type

```{r}
bike_data %>%
  ggplot(aes(x = member_casual, y = trip_duration, fill = member_casual)) +
  geom_boxplot() +
  labs(title = "Boxplot of Trip Durations by Customer Type",
       x = "Customer Type",
       y = "Trip Duration (mins)")
```

## A scatter plot of start and end locations

```{r}
bike_data %>% filter(month =="Jul") %>%
  ggplot(aes(x = start_lng, y = start_lat, color = member_casual)) +
  geom_point(alpha = 0.1) +
  geom_point(aes(x = end_lng, y = end_lat), alpha = 0.1) +
  labs(title = "Scatter Plot of Start and End Locations",
       x = "Longitude",
       y = "Latitude")
```

## A pie chart of customer types

```{r}
bike_data %>%
  count(member_casual) %>%
  ggplot(aes(x = "", y = n, fill = member_casual)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Pie Chart of Customer Types",
       x = NULL,
       y = NULL)
```

## A pie chart by Bike types

```{r}
bike_data %>%
  count(rideable_type) %>%
  ggplot(aes(x = "", y = n, fill = rideable_type)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Pie Chart of Bike Types",
       x = NULL,
       y = NULL)
```

## A density plot of trip durations by customer type

```{r}
bike_data %>%
  ggplot(aes(x = log10(trip_duration), fill = member_casual)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of Trip Durations by Customer Type",
       x = "Trip Duration (mins)log10",
       y = "Density")

```
## A facet plot of bike usage by hour and day of week

```{r}
bike_data %>%
  count(hour, day_of_week) %>%
  ggplot(aes(x = hour, y = n)) +
  geom_area() +
  facet_wrap(~day_of_week) +
  labs(title = "Facet Plot of Bike Usage by Hour and Day of Week",
       x = "Hour",
       y = "Count")
```

## Map of the number of rides for July


